---
title: "BeeProject"
author: "Diane Hu"
date: "November 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library("dplyr")
library("ggplot2")
library("broom")
library("knitr")
library("cowplot")
library("readr")
library("arm")
library("olsrr")
library("car")
```

## Exploratory Data

```{r}
# intial
vHoneyNeonic_v02 <- read_csv("vHoneyNeonic_v02.csv")
summary(vHoneyNeonic_v02)

vHoneyNeonic_v02 <- vHoneyNeonic_v02 %>% mutate(state = as.factor(state), Region = as.factor(Region), StateName = as.factor(StateName))
```
```{r}
glimpse(vHoneyNeonic_v02)
```

Data Clean Up:
```{r}
# need to sample 25% of data so let us sample 200 points
vHoneyNeonic_v02 <- na.omit(vHoneyNeonic_v02) #remove NA data
set.seed(7)

c <- 10**(-8)

# add in indicator variables
#vHoneyNeonic_v02 <- vHoneyNeonic_v02 %>% mutate(in_nACETAMIPRID = as.factor(nACETAMIPRID), in_nCLOTHIANIDIN = as.factor(nCLOTHIANIDIN), in_nIMIDACLOPRID = as.factor(nIMIDACLOPRID), in_nTHIACLOPRID = as.factor(nTHIACLOPRID), in_nTHIAMETHOXAM = as.factor(nTHIAMETHOXAM ))
vHoneyNeonic_v02 <- vHoneyNeonic_v02 %>% mutate(in_nACETAMIPRID = as.numeric(nACETAMIPRID>0), in_nCLOTHIANIDIN = as.numeric(nCLOTHIANIDIN>0), in_nIMIDACLOPRID = as.numeric(nIMIDACLOPRID>0), in_nTHIACLOPRID = as.numeric(nTHIACLOPRID>0), in_nTHIAMETHOXAM = as.numeric(nTHIAMETHOXAM>0))


#shift original values
vHoneyNeonic_v02 <- vHoneyNeonic_v02 %>% mutate(nACETAMIPRID = nACETAMIPRID + c, nCLOTHIANIDIN = nCLOTHIANIDIN + c, nIMIDACLOPRID = nIMIDACLOPRID + c, nTHIACLOPRID = nTHIACLOPRID + c, nTHIAMETHOXAM = c + nTHIAMETHOXAM)

# log and add
vHoneyNeonic_v02 <- vHoneyNeonic_v02 %>% mutate(lognACETAMIPRID = log(nACETAMIPRID), lognCLOTHIANIDIN = log(nCLOTHIANIDIN), lognIMIDACLOPRID = log(nIMIDACLOPRID), lognTHIACLOPRID = log(nTHIACLOPRID),lognTHIAMETHOXAM =log(nTHIAMETHOXAM) )

num <- summarize(vHoneyNeonic_v02,n = n())
honey <- sample_n(vHoneyNeonic_v02, as.integer(num*.25))
```

```{r}
# remove NA items
honey$lognACETAMIPRID[which(!is.finite(honey$lognACETAMIPRID))] = NA
honey$lognCLOTHIANIDIN[which(!is.finite(honey$lognCLOTHIANIDIN))] = NA
honey$lognIMIDACLOPRID[which(!is.finite(honey$lognIMIDACLOPRID))] = NA
honey$lognTHIACLOPRID[which(!is.finite(honey$lognTHIACLOPRID))] = NA
honey$lognTHIAMETHOXAM[which(!is.finite(honey$lognTHIAMETHOXAM))] = NA

honey <- na.omit(honey)
glimpse(honey)
```


## Graphs for yield per colony 
```{r}
yc_nACETAMIPRID <- ggplot(honey, aes(x=nACETAMIPRID, y=yieldpercol)) + geom_point() + 
  ggtitle("Yield per Colony v. nACETAMIPRID")
yc_nCLOTHIANIDIN <- ggplot(honey, aes(x=nCLOTHIANIDIN, y=yieldpercol)) + geom_point() +  
  ggtitle("Yield per Colony v. nCLOTHIANIDIN")
yc_nIMIDACLOPRID <- ggplot(honey, aes(x=nIMIDACLOPRID, y=yieldpercol)) + geom_point() + 
  ggtitle("Yield per Colony v. nIMIDACLOPRID")
yc_nTHIACLOPRID <- ggplot(honey, aes(x=nTHIACLOPRID, y=yieldpercol))  + geom_point() + 
  ggtitle("Yield per Colony v. nTHIACLOPRID")
yc_nTHIAMETHOXAM <- ggplot(honey, aes(x=nTHIAMETHOXAM, y=yieldpercol)) + geom_point() + 
  ggtitle("Yield per Colony v. nTHIAMETHOXAM")
yc_year <- ggplot(honey, aes(x=year, y=yieldpercol)) + geom_point() + ggtitle("Yield per Colony v. Year")
yc_reigon <- ggplot(honey, aes(x=Region, y=yieldpercol)) + geom_boxplot() + ggtitle("Yield per Colony v. Region")

#show plots
yc_nACETAMIPRID
yc_nCLOTHIANIDIN
yc_nIMIDACLOPRID
yc_nTHIACLOPRID
yc_nTHIAMETHOXAM
yc_year
yc_reigon
```


```{r}
yc_lognACETAMIPRID <- ggplot(honey, aes(x=in_nACETAMIPRID*nACETAMIPRID, y=yieldpercol)) + geom_point() +ggtitle("Yield per Colony v. in_nACETAMIPRID")
yc_lognCLOTHIANIDIN <- ggplot(honey, aes(x=in_nCLOTHIANIDIN*nCLOTHIANIDIN, y=yieldpercol)) + geom_point() + ggtitle("Yield per Colony v. in_nCLOTHIANIDIN")
yc_lognIMIDACLOPRID <- ggplot(honey, aes(x=in_nIMIDACLOPRID*nIMIDACLOPRID, y=yieldpercol)) + geom_point() + ggtitle("Yield per Colony v. in_nIMIDACLOPRID")
yc_lognTHIACLOPRID <- ggplot(honey, aes(x=in_nTHIACLOPRID*nTHIACLOPRID, y=yieldpercol))  + geom_point() + ggtitle("Yield per Colony v. in_nTHIACLOPRID")
yc_lognTHIAMETHOXAM <- ggplot(honey, aes(x= in_nTHIAMETHOXAM*nTHIAMETHOXAM, y=yieldpercol)) + geom_point() + ggtitle("Yield per Colony v. in_nTHIAMETHOXAM")

#show plots
yc_lognACETAMIPRID
yc_lognCLOTHIANIDIN
yc_lognIMIDACLOPRID
yc_lognTHIACLOPRID
yc_lognTHIAMETHOXAM
```

indicator (only 1s) * log amount

```{r}
#print these

#get samples with indicator 1
in_1 <- honey %>% filter(in_nACETAMIPRID>0)
in_2 <- honey %>% filter(in_nCLOTHIANIDIN>0)
in_3 <- honey %>% filter(in_nIMIDACLOPRID>0)
in_4 <- honey %>% filter(in_nTHIACLOPRID>0)
in_5 <- honey %>% filter(in_nTHIAMETHOXAM>0) 

yc_lognACETAMIPRID <- ggplot(in_1, aes(x=lognACETAMIPRID, y=yieldpercol)) + geom_point() +ggtitle("Yield per Colony v. lognACETAMIPRID")
yc_lognCLOTHIANIDIN <- ggplot(in_2, aes(x=lognCLOTHIANIDIN, y=yieldpercol)) + geom_point() + ggtitle("Yield per Colony v. lognCLOTHIANIDIN")
yc_lognIMIDACLOPRID <- ggplot(in_3, aes(x=lognIMIDACLOPRID, y=yieldpercol)) + geom_point() + ggtitle("Yield per Colony v. lognIMIDACLOPRID")
yc_lognTHIACLOPRID <- ggplot(in_4, aes(x=lognTHIACLOPRID, y=yieldpercol))  + geom_point() + ggtitle("Yield per Colony v. lognTHIACLOPRID")
yc_lognTHIAMETHOXAM <- ggplot(in_5, aes(x=lognTHIAMETHOXAM, y=yieldpercol)) + geom_point() + ggtitle("Yield per Colony v. lognTHIAMETHOXAM")

#show plots
yc_lognACETAMIPRID
yc_lognCLOTHIANIDIN
yc_lognIMIDACLOPRID
yc_lognTHIACLOPRID
yc_lognTHIAMETHOXAM
```

```{r}
glimpse(ca)
```


CA:

```{r}
glimpse(honey)
```

```{r}
glimpse(ca)
```

```{r}

ca <- vHoneyNeonic_v02 %>% filter(StateName == "California")

#get samples with indicator 1

# we decide to not to move forward with the logged data since removing the infinite values will result in very little observations
year.only <- lm(yieldpercol ~ year, data=ca)
lm.ACETAMIPRID.ca <- lm(yieldpercol ~ year + nACETAMIPRID + nACETAMIPRID*in_nACETAMIPRID + in_nACETAMIPRID, data=ca)
lm.CLOTHIANIDIN.ca <- lm(yieldpercol ~  year+ nCLOTHIANIDIN + nCLOTHIANIDIN*(in_nCLOTHIANIDIN ) + in_nCLOTHIANIDIN , data=ca)
lm.IMIDACLOPRID.ca <- lm(yieldpercol ~  year+nIMIDACLOPRID + nIMIDACLOPRID*(in_nIMIDACLOPRID) + in_nIMIDACLOPRID, data=ca)
lm.THIACLOPRID.ca <- lm(yieldpercol ~  year+nTHIACLOPRID + nTHIACLOPRID*(in_nTHIACLOPRID) + in_nTHIACLOPRID, data=ca)
lm.THIAMETHOXAM.ca <- lm(yieldpercol ~  year+nTHIAMETHOXAM + nTHIAMETHOXAM*(in_nTHIAMETHOXAM) + in_nTHIAMETHOXAM, data=ca)

anova(year.only,lm.ACETAMIPRID.ca)
anova(year.only,lm.CLOTHIANIDIN.ca)
anova(year.only,lm.IMIDACLOPRID.ca)
anova(year.only,lm.THIACLOPRID.ca)
anova(year.only,lm.THIAMETHOXAM.ca)
```
As we see, none of the pestisides are important. We continue with year.only model

```{r}
#print these

ycca_nACETAMIPRID <- ggplot(ca, aes(x=nACETAMIPRID, y=yieldpercol)) + geom_point() +ggtitle("CA: Yield per Colony v. nACETAMIPRID")
ycca_nCLOTHIANIDIN <- ggplot(ca, aes(x=nCLOTHIANIDIN, y=yieldpercol)) + geom_point() + ggtitle("CA: Yield per Colony v. nCLOTHIANIDIN")
ycca_nIMIDACLOPRID <- ggplot(ca, aes(x=nIMIDACLOPRID, y=yieldpercol)) + geom_point() + ggtitle("CA: Yield per Colony v. nIMIDACLOPRID")
ycca_nTHIACLOPRID <- ggplot(ca, aes(x=nTHIACLOPRID, y=yieldpercol))  + geom_point() + ggtitle("CA: Yield per Colony v. nTHIACLOPRID")
ycca_nTHIAMETHOXAM <- ggplot(ca, aes(x=nTHIAMETHOXAM, y=yieldpercol)) + geom_point() + ggtitle("CA: Yield per Colony v. nTHIAMETHOXAM")
ycca_year <- ggplot(ca, aes(x=year, y=yieldpercol)) + geom_point() + ggtitle("CA: Yield per Colony v. Year")

#show plots
ycca_nACETAMIPRID
ycca_nCLOTHIANIDIN
ycca_nIMIDACLOPRID
ycca_nTHIACLOPRID
ycca_nTHIAMETHOXAM
ycca_year

#california w/o 0 indicator variables
in_ycca_nACETAMIPRID <- ggplot(in_1_ca, aes(x=nACETAMIPRID, y=yieldpercol)) + geom_point() +ggtitle("CA: Yield per Colony v. nACETAMIPRID(Indicator)")
in_ycca_nCLOTHIANIDIN <- ggplot(in_2_ca, aes(x=nCLOTHIANIDIN, y=yieldpercol)) + geom_point() + ggtitle("CA: Yield per Colony v. nCLOTHIANIDIN(Indicator)")
in_ycca_nIMIDACLOPRID <- ggplot(in_3_ca, aes(x=nIMIDACLOPRID, y=yieldpercol)) + geom_point() + ggtitle("CA: Yield per Colony v. nIMIDACLOPRID(Indicator)")
in_ycca_nTHIACLOPRID <- ggplot(in_4_ca, aes(x=nTHIACLOPRID, y=yieldpercol))  + geom_point() + ggtitle("CA: Yield per Colony v. nTHIACLOPRID(Indicator)")
in_ycca_nTHIAMETHOXAM <- ggplot(in_5_ca, aes(x=nTHIAMETHOXAM, y=yieldpercol)) + geom_point() + ggtitle("CA: Yield per Colony v. nTHIAMETHOXAM(Indicator)")
in_ycca_year <- ggplot(in_1_ca, aes(x=year, y=yieldpercol)) + geom_point() + ggtitle("CA: Yield per Colony v. Year")

#show plots
in_ycca_nACETAMIPRID
in_ycca_nCLOTHIANIDIN
in_ycca_nIMIDACLOPRID
in_ycca_nTHIACLOPRID
in_ycca_nTHIAMETHOXAM
ycca_year

```

## Model selection 
make the model , do oslrr, trum down , make a model with baseline being the reigon and adding in each pesticeide as we go along.
```{r}
# first model 
full.model <- lm(yieldpercol ~ Region + lognACETAMIPRID + lognCLOTHIANIDIN + lognIMIDACLOPRID + lognTHIACLOPRID + lognTHIAMETHOXAM + (in_nACETAMIPRID) + (in_nCLOTHIANIDIN) + (in_nIMIDACLOPRID) + (in_nTHIACLOPRID) + (in_nTHIAMETHOXAM) + (in_nACETAMIPRID)*lognACETAMIPRID +(in_nCLOTHIANIDIN)*lognCLOTHIANIDIN + (in_nIMIDACLOPRID)*lognIMIDACLOPRID + (in_nTHIACLOPRID)*lognTHIACLOPRID + (in_nTHIAMETHOXAM)*lognTHIAMETHOXAM, data=honey)
tidy(full.model)


reigon.only <- lm(yieldpercol ~ Region, data=honey)
lm.ACETAMIPRID <- lm(yieldpercol ~ Region+ lognACETAMIPRID + lognACETAMIPRID*(in_nACETAMIPRID) + in_nACETAMIPRID, data=honey)
lm.CLOTHIANIDIN <- lm(yieldpercol ~  Region+lognCLOTHIANIDIN + lognCLOTHIANIDIN*(in_nCLOTHIANIDIN) + in_nCLOTHIANIDIN, data=honey)
lm.IMIDACLOPRID <- lm(yieldpercol ~  Region+lognIMIDACLOPRID + lognIMIDACLOPRID*(in_nIMIDACLOPRID) + in_nIMIDACLOPRID, data=honey)
lm.THIACLOPRID <- lm(yieldpercol ~  Region+lognTHIACLOPRID + lognTHIACLOPRID*(in_nTHIACLOPRID) + in_nTHIACLOPRID, data=honey)
lm.THIAMETHOXAM <- lm(yieldpercol ~  Region+lognTHIAMETHOXAM + lognTHIAMETHOXAM*(in_nTHIAMETHOXAM) + in_nTHIAMETHOXAM, data=honey)

anova(reigon.only,lm.ACETAMIPRID)
anova(reigon.only,lm.CLOTHIANIDIN)
anova(reigon.only,lm.IMIDACLOPRID)
anova(reigon.only,lm.THIACLOPRID)
anova(reigon.only,lm.THIAMETHOXAM)

```

nACETAMIPRID (0.01213) and nTHIACLOPRID (0.002801) are significant.

```{r}
# further models
noInteraction.model <- lm(yieldpercol ~ Region + lognTHIACLOPRID + in_nTHIACLOPRID + (in_nTHIACLOPRID*lognTHIACLOPRID) + lognTHIAMETHOXAM + in_nTHIAMETHOXAM + (in_nTHIAMETHOXAM*lognTHIAMETHOXAM), data=honey)
tidy(noInteraction.model)

# consider interactions
interaction.model <- lm(yieldpercol ~ Region + lognTHIACLOPRID + in_nTHIACLOPRID + (in_nTHIACLOPRID*lognTHIACLOPRID) + lognTHIAMETHOXAM + in_nTHIAMETHOXAM + (in_nTHIAMETHOXAM*lognTHIAMETHOXAM) + (lognTHIACLOPRID*Region) +(lognTHIAMETHOXAM*Region), data=honey)
tidy(interaction.model)

# instantiate same interaction model without indicator variables.
interaction.model_0 <- lm(yieldpercol ~ Region + lognTHIACLOPRID + lognTHIAMETHOXAM  + (lognTHIACLOPRID*Region) +(lognTHIAMETHOXAM*Region), data=honey)


```

# assumption verification

Are interactions important?
```{r}
anova(noInteraction.model,interaction.model)

#confirm with the r sqaured 
glance(noInteraction.model)$r.squared
glance(interaction.model)$r.squared

```
P value is small. Therefore we shall continue with interactions included.

```{r}
honey <- honey %>% mutate(Residuals = resid(interaction.model_0))
honey  <- honey %>% mutate(Predict = predict(interaction.model_0))

#resid against explanatory
ggplot(data=honey, aes(x=Region, y=Residuals)) + geom_point() + geom_hline(yintercept=0,color="red") +  labs(title="Residuals vs. Region") + theme(plot.title=element_text(hjust=0.5))

ggplot(data=honey, aes(x=(in_nTHIACLOPRID*lognTHIACLOPRID), y=Residuals)) + geom_point() + geom_hline(yintercept=0,color="red") +  labs(title="Residuals vs. (in_nTHIACLOPRID*lognTHIACLOPRID)") + theme(plot.title=element_text(hjust=0.5))

ggplot(data=honey, aes(x=(in_nTHIAMETHOXAM*lognTHIAMETHOXAM), y=Residuals)) + geom_point() + geom_hline(yintercept=0,color="red") +  labs(title="Residuals vs. (in_nTHIAMETHOXAM*lognTHIAMETHOXAM)") + theme(plot.title=element_text(hjust=0.5))

#resid against explanatory
ggplot(data=honey, aes(x=Predict, y=Residuals)) + geom_point() + geom_hline(yintercept=0,color="red") +  labs(title="Residuals vs. Predict") + theme(plot.title=element_text(hjust=0.5))

```

```{r}
#check normality of residuals
hist(honey$Residuals, main="Histogram of Residuals", xlab="residual") 
ggplot(honey, aes(sample=Residuals))+ stat_qq()+ stat_qq_line() 
```

VIF
```{r}

kable(tidy(vif(interaction.model_0)))
```

Continue checking assumptions:
```{r}
honey <- honey %>% mutate(leverage = hatvalues(interaction.model),
cooks = cooks.distance(interaction.model_0), 
stand.resid = rstandard(interaction.model_0), 
obs.num = row_number())

#Leverage
ggplot(data=honey, aes(x=obs.num,y=leverage)) + 
geom_point(alpha=0.5) + 
geom_hline(yintercept=0.1,color="red")+
labs(x="Observation Number",y="Leverage",title="Leverage")

#Cooks
ggplot(data=honey, aes(x=obs.num,y=cooks)) + 
geom_point() + 
geom_hline(yintercept=1,color="red")+
labs(x="Observation Number",y="Cook's Distance",title="Cook's Distance")

# r sq
glance(interaction.model_0)$r.squared
glance(interaction.model_0)$adj.r.squared
```


Let us look at the second model for CA only (yclm_2):
```{r}
ca.model <- year.only #rename
ca <- ca %>% mutate(Residuals_2 = resid(ca.model))
ca  <- ca %>% mutate(Predict_2 = predict(ca.model))


#resid against explanatory
ggplot(data=ca, aes(x=year, y=Residuals_2)) + geom_point() + geom_hline(yintercept=0,color="red") +  labs(title="Residuals vs. year") + theme(plot.title=element_text(hjust=0.5))

#resid against explanatory
ggplot(data=ca, aes(x=Predict_2, y=Residuals_2)) + geom_point() + geom_hline(yintercept=0,color="red") +  labs(title="Residuals vs. Predict_2") + theme(plot.title=element_text(hjust=0.5))

```

```{r}
#check normality of residuals
hist(ca$Residuals_2, main="Histogram of Residuals", xlab="residual") 
ggplot(ca, aes(sample=Residuals_2))+ stat_qq()+ stat_qq_line() 
```


Continue checking assumptions:
```{r}
ca <- ca %>% mutate(leverage = hatvalues(ca.model),
cooks = cooks.distance(ca.model), 
stand.resid = rstandard(ca.model), 
obs.num = row_number())

#Leverage
ggplot(data=ca, aes(x=obs.num,y=leverage)) + 
geom_point(alpha=0.5) + 
geom_hline(yintercept=0.1,color="red")+
labs(x="Observation Number",y="Leverage",title="Leverage")

#Cooks
ggplot(data=ca, aes(x=obs.num,y=cooks)) + 
geom_point() + 
geom_hline(yintercept=1,color="red")+
labs(x="Observation Number",y="Cook's Distance",title="Cook's Distance")

# r sq
glance(ca.model)$r.squared
glance(ca.model)$adj.r.squared
```

Prediction of Future Values 
```{r}
# for pesticides generally

interaction.model_0 <- lm(yieldpercol ~ Region + lognTHIACLOPRID + lognTHIAMETHOXAM  + (lognTHIACLOPRID*Region) +(lognTHIAMETHOXAM*Region), data=honey)

predict_model <- data.frame(Reigon = c("Midwest","Northeast", "South","West"),
                            )
predict.lm(interaction.model_0,newdata=predict_model,interval="prediction")


# for california
predict_cal <- data.frame(year = c(2050,2100))
predict.lm(ca.model,newdata=predict_cal,interval="prediction")
```
